---
title: "A Look at the Endangered & Threatened Animals Dataset"
author: "Ana De Las Alas, Samuel Lee, Ryan Leung, Joshua Lansing"
date: "May 24, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(knitr)
source("scripts/summary.R")
source("scripts/aggregate_table.R")
source("scripts/chart1_delasalas.R")
source("scripts/chart2_lee.R")
source("scripts/chart3_leung.R")
```

```{r dataset, echo = FALSE}
# Reads dataset on endangered species
species_dataset <- read.csv("data/us_endangered_species.csv",
                            stringsAsFactors = FALSE)
```

## Introduction
Our team is focusing on data about _endangered and threatened_ species in the U.S. Our dataset was taken from the [Environmental Convervation Online System](https://ecos.fws.gov/ecp/report/table/critical-habitat.html?fbclid=IwAR34iloupBtcBAHfH7t8uU0WqRWJyGDKggJShn-J1mSbNVl_Zf_7qYP8kFI). The dataset includes data about specific endangered species and their respective attributes, such as the **region** of the U.S. they are found in, their **species group**, and their **critical habitat type**. With this domain, we can investigate questions like which species group contains the most endangered species, which region is likely to contain certain species groups, and what are the patterns in how many acres are allotted to create critical habitats for these species. The following report will highlight summary information about the dataset as well as three visualizations that convey key points about the data. 

## Summary Information 

```{r summary list, echo = FALSE}
list <- summary_information(species_dataset)
```
Within the dataset, there are `r list$endangered` total **endangered** species and `r list$threatened` total **threatened** species documented. The percentage of species that are endangered is `r list$endangered_percentage`%. Our dataset shows which _species group_ each species belongs to. Based on our calculations, we found that there were `r list$mammals` mammals documented in the dataset, which means that `r list$mammal_percentage`% of the species from the data are mammals. The organization separated the U.S. into `r list$regions` regions so that readers can better understand where each species in the dataset comes from. This information helps our team recognize that there are more endangered species than threatened species, that mammals make up a small percentage of endangered species, and that the location of the species is dependent on region rather than state or city. 

## Aggregate Table
This table shows the distribution of endangered/threatened animals across the
various regions in the U.S. Some notable information from this table that is show
is that the **Pacific**, **Southeast** and **Pacific Southwest** regions have the
**highest** counts of endangered/threatened animals. In contrast, the **Southeast** and
**Alaska** regions have the **lowest** counts.
```{r aggregate table, echo = FALSE}
summary_table(endangered_animals)
```

## Chart One: Boxplot 
I included this boxplot to show the **distribution** of data regarding the area of critical habitats meant to sustain endangered species. The areas are defined by the number of acres that a critical habitat takes up. This boxplot was intended to convey the medians, quartile ranges, and outliers of the values in the dataset. We are able to identify that the species group with the highest median is _mammals_, that critical habitats for crustaceans tend to take up, on average, a smaller area, and that critical habitats for reptiles have the largest range in acres. This boxplot helps us answer questions like how much space is allotted to preserve a critical habitat for a certain species group. 

```{r boxplot, warning = FALSE, echo = FALSE}
boxplot(acre_data)
```

## Chart Two: Bar Graph
This chart was made to show the contrast of *endangered species* in different regions. We decided to use a bar graph for this because it would help us focus in on particularly the endangered species and would help us easily see the difference in quantity of endangered species according to region. **Region 1** has the most amount of endangered species, while **Region 3** has the least amount. **Region 4** has a similar amount as **Region 8** both under *100 species*.

```{r pressure, echo=FALSE}
regions_plot(tally_lead_region)
```


## Chart Three: Lollipop Graph  
The purpose of this chart is to display the number of endangered and threatened species within each major U.S. region. From the chart, it can be seen that the **West Coast** has the most number of endangered and threatened, whereas the **East Coast** has the least.

```{r chart, warning=FALSE, message=FALSE, echo=FALSE}
plot_leung(endangered_v5)
```
